### 2014-10-20 ApTranscriptome_TR.Rmd

# Pearson Chi-square test
chi1 <- chisq.test(Ap.type.table)
chi1

# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Acar = Ap.response.type[ , 'Ar.type'], Apic = Ap.response.type[ , 'A22.type'])
# reorder
tt2 <- type.table[c("Low", "Intermediate", "High", "Bimodal", "NotResp"), c("Low", "Intermediate", "High", "Bimodal", "NotResp")]

@@

Table with 'Total Transcripts' category

```{r total_table}

```

@@

```{r mh_plot}
# overall mean for each class
rs <- rowSums(tt2)
cs <- colSums(tt2)
(gm <- (rs + cs) / sum(tt2 * 2))

@@

md <- melt(Ec.cells)

@@

# grid.arrange
tiff("results/constitutive_vs_inducibility.tiff", compression = "lzw", bg = "white")
grid.arrange(gg5, gg6, nrow = 1)
dev.off()

### plot constitutive expression in one species vs the other
gg7 <- ggplot(Ap.response.type2.up, aes(x = log(Ar.opt), y = log(A22.opt), colour = A22.type)) + 
  geom_point() + 
  geom_smooth() +
  labs(x="log(consitutive expression of A. carolinensis)", y = "log(constitutive expression in A. picea)") +
  theme(legend.position = "none")  
gg7

@@

@@

gg9 <- ggplot(Ap.response.type2.A22.high, aes(x = log(Ar.opt), y = log(A22.opt), colour = Ar.type)) + 
  geom_point() + 
  geom_smooth() +
  labs(x="log(consitutive expression of A. carolinensis)", y = "log(Degree of inducibility in A. picea)") +
  theme(legend.position = "none")  
gg9

```

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(TPM.dt.sub, RxNpval, sigtable, chi1, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, Ap.T_on, Ap.T_off, hsp_responsive, hsp_all, mh.test, mh_plot, Ap.BP.resultParentChild, resp.stress.responsive.count, T_on.high.ttest, T_on.low.ttest, T_off.int.high.ttest, T_off.int.low.ttest, Ap.gsea.intersect, Ar.int.A22.bim.gsea, Ar.int.A22.low.gsea, Ar.int.A22.lowbim.gsea, quadratic.lms, resp.TPM.dt.sub.pred, md, file = paste("doc/ms_", Sys.Date(), ".RData", sep = ""))
```

>>>>

# test if the total number of responsive transcripts differs between species 
resp.table <- matrix(nrow=2, ncol=2, dimnames = list(c("R", "NR"), c("Ap", "Ac")))
resp.table[1,1] <- sum(A22.type.table[which(names(A22.type.table) != "NotResp")])
resp.table[2,1] <- nrow(annotation.table) - sum(A22.type.table[which(names(A22.type.table) != "NotResp")])
resp.table[1,2] <- sum(Ar.type.table[which(names(Ar.type.table) != "NotResp")])
resp.table[2,2] <- nrow(annotation.table) - sum(Ar.type.table[which(names(Ar.type.table) != "NotResp")])
resp.table

chi1 <- chisq.test(resp.table)

# test if the number of transcripts in each category differs from null expectations using Pearson Chi-square test
chi2 <- chisq.test(Ap.type.table)
chi2
```

The total number of responsive transcripts does not differ between species, though their groupings into expression categories does differ.

```{r mosaic_plot}
# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Acar = Ap.response.type[ , 'Ar.type'], Apic = Ap.response.type[ , 'A22.type'])
# reorder
tt2 <- type.table[c("Low", "Intermediate", "High", "Bimodal", "NotResp"), c("Low", "Intermediate", "High", "Bimodal", "NotResp")]

@@

@@

To identify specific cells that deviate from null expectation, calculate expected observations assuming that marginal frequencies equal the overall frequency for each expression category, then determine which cells have deviations between observed and expected that are greater than the overall X^2^ statistic.

```{r chi_posthoc}
# overall mean for each class
rs <- rowSums(tt2)
cs <- colSums(tt2)
(gm <- (rs + cs) / sum(tt2 * 2))

@@

# which of these are significantly great than expected chi-squared?
```

To visualize this, make a heatmap that shows deviation from expected count.

```{r mh_plot}
md <- melt(Ec.cells)

@@

# grid.arrange
png("results/constitutive_vs_inducibility.png")
grid.arrange(gg5, gg6, nrow = 1)
dev.off()

### plot constitutive expression in one species vs the other
gg7 <- ggplot(Ap.response.type2.up, aes(x = log(Ar.opt), y = log(A22.opt), colour = A22.type)) + 
  geom_point() + 
  geom_smooth() +
  labs(x="log(consitutive expression of A. carolinensis)", y = "log(constitutive expression in A. picea)") +
  theme(legend.position = "none")  
gg7 <- gg7 + geom_abline(intercept=0, slope=1)

png("results/constitutive_comparison.png")
gg7
dev.off()

@@

png("results/upregulation_comparison.png")
gg8
dev.off()

@@

gg9 <- ggplot(Ap.response.type2.A22.high, aes(x = log(Ar.opt), y = log(A22.opt), colour = Ar.type)) + 
  geom_point() + 
  geom_smooth() +
  labs(x="log(consitutive expression of A. carolinensis)", y = "log(Degree of inducibility in A. picea)") +
  theme(legend.position = "none")  
gg9 + geom_abline(intercept=0, slope=1)

png("results/constitutive_vs_inducibility_A22high.png")
gg9
dev.off()
```

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(TPM.dt.sub, RxNpval, sigtable, chi1, chi2, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, hsp_responsive, hsp_all, mh.test, mh_plot, Ap.BP.resultParentChild, resp.stress.responsive.count, T_on.high.ttest, T_on.low.ttest, T_off.int.high.ttest, T_off.int.low.ttest, Ap.gsea.intersect, Ar.int.A22.bim.gsea, Ar.int.A22.low.gsea, Ar.int.A22.lowbim.gsea, quadratic.lms, resp.TPM.dt.sub.pred, md, file = paste("doc/ms_", Sys.Date(), ".RData", sep = ""))
```
